Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-08-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add some to text and references

Good Characteristics:
- Measureable at the county/sub-county level
- Internally consistant (either within a study or for a particular storm)
+ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
- Flooding-- Most associated with death; property damage
- Precipitation
+ Gauge data if available
- Wind (potentially weighted based on relation to the storm track), though wind alone has proven not to be sufficient in identifying the most affected region in some cases (e.g., Hurricane Andrew in 1992)[@Fronstin1994]. Therefore wind alone is not sufficient in all cases (as is commonly done using the Staffir-Simpson scale).

Bad Characteristics:
- Distance alone
- Wind alone
- Metrics that cannot be widely measured at the county or sub-county level (i.e., stress)

Hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used? We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale). These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016; @Fronstin1994]. FEMA declaration is largely subjective, requiring state governors and/or tribal chiefs to submit a proposal for government aid on the basis that local resources are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid. As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane; GBA-- let's think about why that might have had "false positives". I think that would be an interesting point to include here.]. Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015]. Such missclassification can have important consequences in assessments of the public health and economic impacts of tropical storms.

@@

If exposure to tropical storms over multiple storms and years can be assessed, these exposure datasets can be joined with other time series to explore the impacts of tropical storms. For example, daily counts of human health outcomes in environmental epidemiology studies are often available aggregated at the county level, and such data has often been paired with time series of environmental exposures (e.g., air pollution, temperature) to determine associated risks. Databases of storm impacts, including direct deaths [@Rappaport2000], ..., and ..., are often aggregated at the county level. [Let's see if we can find other examples of county-level outcomes that researchers might want to pair with hurricane exposure. For example, property damage, crop damage, injuries, ... These are all given in NOAA's Storm Events database, but let's also see if we can find studies that have analyzed these outcomes in relation to tropical storms, if we can.]

@@

A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000]. One study of US hurricanes between 1925 and 1995 found that property damage associated with a storm was strongly associated with the intensity of the storm, with most damage occuring during category 3, 4, or 5 storms, even though storms of these categories made up under a quarter of all hurricanes in the study [@Pielke1998]. Wind-based hazards are also closely associated with power outages [citation], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006]. Direct fatalities from hurricanes tend to increase with increasing wind exposure from the storm [@Czajkowski2011].

>>>>

Good Characteristics:
- Measureable at the county/sub-county level
- Internally consistant (either within a study or for a particular storm)
+ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
- Flooding-- Most associated with death; property damage
- Precipitation
+ Gauge data if available
- Wind (potentially weighted based on relation to the storm track), though wind alone has proven not to be sufficient in identifying the most affected region in some cases (e.g., Hurricane Andrew in 1992)[@Fronstin1994]. Therefore wind alone is not sufficient in all cases (as is commonly done using the Staffir-Simpson scale).
- Have a metric with finer resolution than the community level (i.d., zip codes or neighborehoods)
+ @Fronstin1994 stated that neighborehoods just 8 blocks away from some of the most impacted areas during Hurricane Andrew (1992) recieved far less damage 
+ Not necessarily possible when guage data may not be available at such a resolution
Bad Characteristics:
- Distance alone
- Wind alone
- Metrics that cannot be widely measured at the county or sub-county level (i.e., stress)

Hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used? We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath based on a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale). These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016; @Fronstin1994]. FEMA declaration is largely subjective, requiring state governors and/or tribal chiefs to submit a proposal for government aid on the basis that local resources are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid. (FEMA is likely less objective given societal effects (politics) [@Kunkel1999]) As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane; GBA-- let's think about why that might have had "false positives". I think that would be an interesting point to include here.]. Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015]. Such missclassification can have important consequences in assessments of the public health and economic impacts of tropical storms.

@@

If exposure to tropical storms over multiple storms and years can be assessed, these exposure datasets can be joined with other time series to explore the impacts of tropical storms. For example, daily counts of human health outcomes in environmental epidemiology studies are often available aggregated at the county level, and such data has often been paired with time series of environmental exposures (e.g., air pollution, temperature) to determine associated risks. Databases of storm impacts, including direct deaths [@Rappaport2000], ..., and ..., are often aggregated at the county level. [Let's see if we can find other examples of county-level outcomes that researchers might want to pair with hurricane exposure. For example, property damage, crop damage, injuries, ... These are all given in NOAA's Storm Events database, but let's also see if we can find studies that have analyzed these outcomes in relation to tropical storms, if we can.] Review has shown the hurricanes are associated with major societal impacts.

Though the loss of life due to hurricanes may have decreased since the mid-1900s, trends show a large increase in property damage [@Kunkel1999] in spite of a statistically significant decrease in both the frequency and intensity of tropical storms since then [@Landsea1996]. Such a trend is likely attributable to increased population growth in coastal communities [@Pielke1997].

@@

A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000]. One study of US hurricanes between 1925 and 1995 found that property damage associated with a storm was strongly associated with the intensity of the storm, with most damage occuring during category 3, 4, or 5 storms, even though storms of these categories made up under a quarter of all hurricanes in the study [@Pielke1998]. Wind-based hazards are also closely associated with power outages (both specific to hurricane related hazards; do not specify wind-based or otherwise) [@Liu2005; @Han2009], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006]. Direct fatalities from hurricanes tend to increase with increasing wind exposure from the storm [@Czajkowski2011].
